---
title: "Telemundo on Peacock — Executive Summary"
format:
  revealjs:
    theme: simple
    slide-number: true
    transition: fade
    css: styles.css
    center: true
    width: 1280
    height: 720
    auto-stretch: false
execute:
  echo: false
  message: false
  warning: false
---


```{r include=FALSE, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(scales)

```



## Executive Summary

### What Defines Success for Telemundo on Peacock?

Success is driven by:

- Early Spanish-language engagement (Week 1)
- High-intensity Spanish viewing behavior
- Sustained subscription durability

Spanish-dominant, high-engagement users drive disproportionate platform value.

## Early Spanish Exposure = Durable Subscribers

- **95% retention vs 89% baseline**
- **+6pp 30-day lift**

**Implication:** Prioritize Spanish-first onboarding in Week 1.

```{r echo=FALSE, fig.height=3.8, fig.width=10, out.width="70%"}
library(readr)
library(dplyr)
library(ggplot2)
library(scales)

kpi <- read_csv("data/clean_retention_early_group_30d.csv") %>%
  mutate(
    retention = retention_30d * 100,
    group = early_group
  ) %>%
  arrange(desc(retention)) %>%
  mutate(group = factor(group, levels = group))

group_colors <- c(
  "Spanish Early Only" = "#1f77b4",
  "Both Languages Early" = "#4c78a8",
  "No Early Engagement" = "#9aa0a6",
  "English Early Only" = "#c0c4c9"
)

base_theme <- theme_minimal(base_size = 20) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    plot.title = element_text(face = "bold", size = 24),
    plot.title.position = "plot",
    legend.position = "none"
  )

# Baseline reference line
baseline <- kpi$retention[kpi$group == "No Early Engagement"][1]

ggplot(kpi, aes(x = group, y = retention, fill = group)) +
  geom_col(width = 0.55) +
  
  # Thinner dashed baseline
  geom_hline(
    yintercept = baseline,
    linetype = "dashed",
    linewidth = 0.6,   # thinner line
    color = "black"
  ) +
  
  geom_text(
    aes(label = paste0(round(retention, 1), "%")),
    vjust = 1.3,
    color = "white",
    size = 6,
    fontface = "bold"
  ) +
  
  # Baseline label moved to right side and above line
annotate(
  "text",
  x = length(levels(kpi$group)),  # far right category
  y = baseline + 7,               # slightly above dashed line
  label = paste0("Baseline: ", round(baseline, 1), "%"),
  hjust = 1,                      # right-align text
  size = 5,
  fontface = "bold",
  color = "black"
)+
  
  scale_fill_manual(values = group_colors) +
  scale_y_continuous(
    limits = c(0, 100),
    breaks = seq(0, 100, 20),
    expand = c(0, 0)
  ) +
  labs(
    title = "30-Day Retention by Early Language Exposure",
    y = NULL
  ) +
  base_theme +
  theme(
    axis.text.x = element_text(size = 14, angle = 15, hjust = 1),
    plot.margin = margin(20, 20, 20, 20)
  )
```

## Engagement Concentration Drives Economic Value

- **38% of users drive 63% of total minutes**
- High-intensity Spanish viewers power platform economics

**Strategic Priority:** Protect and deepen this segment.
```{r echo=FALSE, fig.height=3.8, fig.width=10, out.width="70%"}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(readr)

cluster_profile <- read_csv("data/cluster_profile.csv") %>%
  mutate(
    pct_users = as.numeric(sub("%","", pct_users)) / 100,
    pct_total_minutes = as.numeric(sub("%","", pct_total_minutes)) / 100
  ) %>%
  mutate(
    cluster_name = case_when(
      cluster_name == "English-First / At-Risk" ~ "Spanish Power Users",
      cluster_name == "Spanish Power Users" ~ "Spanish-Light Users",
      cluster_name == "Spanish-Light Users" ~ "Low-Engagement Users"
    )
  ) %>%
  mutate(
    cluster_name = factor(cluster_name, levels = c(
      "Spanish Power Users",
      "Spanish-Light Users",
      "Low-Engagement Users"
    ))
  )

cluster_profile <- cluster_profile %>%
  mutate(gap = pct_total_minutes - pct_users)

# Identify segment with largest positive gap
top_gap <- cluster_profile %>%
  arrange(desc(gap)) %>%
  slice(1)

ggplot(cluster_profile, aes(y = cluster_name)) +
  
  geom_segment(
    aes(x = pct_users, xend = pct_total_minutes, yend = cluster_name),
    linewidth = 4,
    color = "#c0c4c9"
  ) +
  
  geom_point(aes(x = pct_users), size = 6, color = "#9aa0a6") +
  geom_point(aes(x = pct_total_minutes), size = 6, color = "#1f77b4") +
  
  # Gap annotation
  annotate(
    "text",
    x = top_gap$pct_total_minutes + 0.02,
    y = top_gap$cluster_name,
    label = paste0("+", round(top_gap$gap * 100, 1), "pp gap"),
    fontface = "bold",
    size = 5,
    hjust = 0
  ) +
  
  scale_x_continuous(
    labels = scales::percent_format(accuracy = 1),
    limits = c(0, max(cluster_profile$pct_total_minutes) + 0.1)
  ) +
  
  labs(
    title = "User Share vs Total Viewing Share",
    subtitle = "Blue = Share of Total Minutes | Gray = Share of Users",
    x = NULL,
    y = NULL
  ) +
  
  theme_minimal(base_size = 20) +
  theme(
    plot.title = element_text(face = "bold", size = 26),
    plot.subtitle = element_text(size = 18),
    panel.grid.minor = element_blank()
  )



```

## Early Spanish Engagement Reduces Churn

- **5.2% vs 11.2% 30-day churn**
- **−6pp absolute reduction**
- **~54% lower churn**

**Action:** Expand early Spanish exposure in onboarding.
```{r echo=FALSE, fig.height=3.8, fig.width=10, out.width="70%"}
library(readr)
library(dplyr)
library(ggplot2)
library(scales)

churn_kpi <- read_csv("data/churn_spanish_early_30d.csv") %>%
  mutate(
    group = ifelse(spanish_early == 1, "Spanish Early (≤7d)", "Not Spanish Early"),
    churn = churn_30d * 100
  ) %>%
  select(group, users, churn) %>%
  mutate(group = factor(group, levels = c("Spanish Early (≤7d)", "Not Spanish Early")))

group_colors <- c(
  "Spanish Early (≤7d)" = "#1f77b4",
  "Not Spanish Early"   = "#9aa0a6"
)

baseline <- churn_kpi$churn[churn_kpi$group == "Not Spanish Early"][1]

ggplot(churn_kpi, aes(x = group, y = churn, fill = group)) +
  geom_col(width = 0.55) +
  
  # Baseline (Not Spanish Early)
  geom_hline(
    yintercept = baseline,
    linetype = "dashed",
    linewidth = 0.6,
    color = "black"
  ) +
  
  # % labels above bars
  geom_text(
    aes(label = paste0(round(churn, 1), "%")),
    vjust = -0.6,
    color = "black",
    size = 6,
    fontface = "bold"
  ) +
  
  # Baseline label on right side, slightly above line

  
  scale_fill_manual(values = group_colors) +
  scale_y_continuous(
    limits = c(0, max(churn_kpi$churn) + 3),
    breaks = seq(0, 20, 5),
    expand = c(0, 0)
  ) +
  labs(
    title = "30-Day Churn by Early Spanish Exposure",
    y = NULL
  ) +
  
  theme_minimal(base_size = 20) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(face = "bold", size = 16),
    plot.title = element_text(face = "bold", size = 24),
    plot.title.position = "plot",
    legend.position = "none",
    plot.margin = margin(20, 20, 20, 20)
  )
```



## What We Should Prioritize Next Quarter

### 1. Run Spanish-First Onboarding A/B Test  
Increase Week-1 visibility of Telemundo content.

### 2. Optimize Merchandising for New Users  
Promote high-retention franchises and starter bundles.

### 3. Measure Incrementality  
Control for acquisition channel and engagement intensity.
