---
title: "Telemundo on Peacock — Executive Summary"
format:
  revealjs:
    theme: simple
    slide-number: true
    transition: fade
    css: styles.css
    center: true
    width: 1280
    height: 720
    auto-stretch: false
execute:
  echo: false
  message: false
  warning: false
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)

kpi <- data.frame(
  group = c("Spanish Early (≤7d)", "Not Spanish Early"),
  retention = c(94.8, 88.8),
  churn = c(20.0, 41.1)
)

# Color mapping (Spanish Early vs Not)
kpi$group <- factor(kpi$group, levels = c("Spanish Early (≤7d)", "Not Spanish Early"))
group_colors <- c("Spanish Early (≤7d)" = "#1f77b4",  # blue
                  "Not Spanish Early"   = "#9aa0a6")  # soft gray

base_theme <- theme_minimal(base_size = 20) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    plot.title = element_text(face = "bold", size = 24),
    plot.title.position = "plot",
    legend.position = "none"
  )

```

## What Defines Success for Telemundo on Peacock

### Success = Subscription Durability

Telemundo content must:

- Drive early engagement (first 7 days)
- Increase **30-day retention**
- Reduce **churn**
- Build durable viewing habit

Success is not viewership alone —  
it is **measurable impact on retention and lifetime value.**


## Most Important Insight

### Early Spanish Engagement Predicts Retention

```{r echo=FALSE, fig.height=3.8, fig.width=10, out.width="75%"}
ggplot(kpi, aes(x = group, y = retention, fill = group)) +
  geom_col(width = 0.55) +
  geom_text(
    aes(label = paste0(retention, "%")),
    vjust = 1.3,              # pushes text down inside the bar
    color = "white",          # makes text readable on dark blue
    size = 6,
    fontface = "bold"
  ) +
  scale_fill_manual(values = group_colors) +
  scale_y_continuous(
    limits = c(0, 100),
    breaks = seq(0,100,20),
    expand = c(0,0)
  ) +
  labs(
    title = "30-Day Retention",
    y = NULL
  ) +
  base_theme +
  theme(
    axis.text.x = element_text(size = 14),
    plot.margin = margin(20, 20, 20, 20)
  )
```
- **94.8%** vs **88.8%** retention
- **+6pp retention lift** for Spanish-early viewers


## Churn Reduction

### Spanish Early Engagement = Lower Churn

Spanish-language content:
```{r echo=FALSE, fig.height=3.8, fig.width=10, out.width="75%"}
ggplot(kpi, aes(x = group, y = churn, fill = group)) +
  geom_col(width = 0.55) +
  geom_text(
    aes(label = paste0(churn, "%")),
    vjust = -0.6,
    size = 6,
    fontface = "bold"
  ) +
  scale_fill_manual(values = group_colors) +
  scale_y_continuous(limits = c(0, 100), breaks = seq(0,100,20), expand = c(0,0)) +
  labs(
    title = "Churn Rate",
    y = NULL
  ) +
  base_theme +
  theme(
    axis.text.x = element_text(size = 14),
    plot.margin = margin(20, 20, 20, 20)
  )
```
- **20.0%** vs **41.1%** churn
- **~50% lower churn** among Spanish-early users
- Pattern holds across preferred language and acquisition channel

## What We Should Prioritize Next Quarter

### Action Plan

**1. Run Spanish-first onboarding A/B test**  
Increase visibility of Telemundo content in first-session UX.

**2. Optimize merchandising for new users**  
Surface high-retention franchises and starter packs.

**3. Measure incrementality**  
Quantify retention lift controlling for channel and language.

**Objective:** Maximize durable subscription value through early Spanish engagement.



